Online {\IterComp}

### 2017-09-16 infra.tex

Figure~\ref{fig:infra-diagram} shows an overview of how our online {\itercomp} mechanism works. On the developer's side, the program is
pre-compiled to IR without optimizations and is instrumented to measure the amount of work performed and the wall-clock time. On the user's
side, we select an optimization sequence and then compile the instrumented IR to binary using this sequence. Next time the user runs the
application, we measure the work and the execution time. We accumulate multiple such measurements until we can estimate the efficiency
metric P of this optimization sequence with high confidence. Finally, we use this information to help us select the next optimization
sequence.

This work focuses on the highlighted component of Figure~\ref{fig:infra-diagram}, the \textit{Work Instrumentation}. It is performed once
and offline, right after the IR is generated. Its purpose is to measure the amount of work performed. Conceptually, we use the cost model of
Section~\ref{sec:metric} to estimate a cost for each basic block of the IR and then add instrumentation to increase a global work counter
by that cost every time a basic block is executed. The measured work represents the computational effort of the unoptimised program and it
should depend only on the program and the input, not the applied optimizations. For that reason, we instrument the unoptimized IR before any optimizations modify it. We present a detailed description of work
instrumentation and how to keep it lightweight in Section~\ref{sec:prof}.

>>>>

Figure~\ref{fig:infra-diagram} shows an overview of how our online {\itercomp} mechanism works. On the developer's side, the program is
pre-compiled to IR without optimizations and is instrumented to measure the amount of work performed and the wall-clock time. On the user's
side, we select an optimization sequence and then compile the instrumented IR to binary using this sequence. Next time the user runs the
application, we measure the work and the execution time. We accumulate multiple such measurements, executing each input only once, until we
can estimate the efficiency metric P of this optimization sequence with high confidence. Finally, we use this information to help us select
the next optimization sequence.

This work focuses on the highlighted component of Figure~\ref{fig:infra-diagram}, the \textit{Work Instrumentation}. It is performed once
and offline, right after the IR is generated. Its purpose is to measure the amount of work performed. Conceptually, we use the cost model of
Section~\ref{sec:metric} to estimate a cost for each basic block of the IR and then add instrumentation to increase a global work counter
by that cost every time a basic block is executed. The measured work represents the computational effort of the unoptimized program and it
should depend only on the program and the input, not the applied optimizations. For that reason, we instrument the unoptimized IR before any optimizations modify it. We present a detailed description of work
instrumentation and how to keep it lightweight in Section~\ref{sec:prof}.

### 2017-09-16 related.tex

\subsection{{\IterComp}}
Various adaptive compilation techniques have been proposed for finding near optimal compilation
settings~\cite{agakov06,kulkarni04,stephenson03}. These approaches evaluate the performance of candidate compiler options through repeated
executions using the same input. All these approaches implicitly assume the found compiler settings work for other unseen datasets.
Chen~\etal~\cite{chen10,chen12a} show that this assumption is not true, and the best optimization options can differ across inputs. Still,
in most cases it is possible to find a combination of optimizations that produces good results across all inputs, by using a few tens of
inputs to evaluate our optimizations on. While a positive result, this is still impractical. Producing tens of realistic input sets might
be difficult. Even if not, each combination of optimization options has to be evaluated on every input set, increasing the time \itercomp
takes by at least an order of magnitude.

Recent research~\cite{chen12b,fang15} has used real data for performing online {\itercomp} on distributed data center applications. Each
worker receives a subset of the input dataset to evaluate a small set of optimization settings on. The best such setting from each round
is used for subsequent executions of the same code. By testing new settings and re-evaluating old ones on new datasets this approach makes
it more likely to select an optimization setting that works well across inputs. Like in~\cite{chen10} every setting is tested on the same
set of inputs, wasting time and energy. This approach only works well with MapReduce-like workloads, since it relies on the MapReduce
framework for repeating the same computation multiple times without side-effects or corrupting the input.

>>>>

\subsection{{\IterComp}}
Various adaptive compilation techniques have been proposed for finding near optimal compilation
settings~\cite{agakov06,kulkarni04,stephenson03}. These approaches evaluate the performance of candidate compiler options through repeated
executions using the same input. All these approaches implicitly assume the best-performing compiler settings found from a small set of
inputs will work well on unseen datasets. Chen~\etal~\cite{chen10,chen12a} show that the best optimization options can differ across
inputs. Their results suggest that, in most cases it is possible to find a combination of optimizations that produces good results across
all inputs, by using a few tens of inputs to evaluate our optimizations on. While a positive result, this is still impractical. Producing
tens of realistic input sets might be difficult. Even if not, each combination of optimization options has to be evaluated on every input
set, increasing the time \itercomp takes by at least an order of magnitude.
%=======
%Multiple adaptive compilation techniques have been proposed for finding a near optimal combination of optimization
%settings~\cite{agakov06,hoste08,kulkarni04,stephenson03} by probabilistically searching the space of possible combinations. But with
%all of them there is uncertainty about whether the quality of their results depends on the data set used. Chen~\etal~\cite{chen10,chen12a}
%examined this issue and showed that indeed the best optimization options differ across inputs. Still, in most cases it is possible to find
%an optimization sequence that produces good results across all inputs, by evaluating the optimizations on a representative set inputs.
%While a positive result, this is still impractical. Producing a set of realistic inputs might be difficult. Even if not, each
%optimization sequence has to be evaluated on every input, increasing the time \itercomp takes by at least an order of
%magnitude.
%>>>>>>> origin/master

Recent research~\cite{chen12b,fang15} has used real data for performing {\itercomp} on distributed data center applications. Each
worker receives a subset of the input dataset to evaluate a small set of optimization settings on. The best such setting from each round
is used for subsequent executions of the same code. By testing new settings and re-evaluating old ones on new datasets this approach makes
it more likely to select an optimization setting that works well across inputs. Like in~\cite{chen10} every setting is tested on the same
set of inputs, wasting time and energy. This approach only works well with MapReduce-like workloads, since it relies on the MapReduce
framework for repeating the same computation multiple times without side-effects or corrupting the input.
